\section{Implementation}
\subsection{Hardware}
For the hardware implementation of TinyNet, we developed compact routers powered by the ATSAMS70 microcontroller from Atmel/Microchip, which operates at a core clock speed of 300MHz. The router schematic and board file are illustrated in Figs. \ref{fig:router-schematic} and \ref{fig:router-board-file} in the appendix. While this microcontroller is capable of supporting UART transmission speeds up to 18.75 Mbps, we observed that a reduction in bitrate to 3.125 Mbps was necessary to achieve stable and reliable data transmission. At higher bitrates, the inter-byte arrival interval becomes shorter than the UART interrupt service routine latency on the ATSAMS70, causing receive buffer overruns; the 3.125 Mbps setting provides sufficient margin for the 300 MHz core to service each incoming byte interrupt before the next byte arrives.

To enhance the system's resilience to electromagnetic interference (EMI), we incorporated the ISL3177 differential driver into the UART lines. Although this measure wasn't strictly required for our initial testing environment, it is a crucial step towards preparing TinyNet for operation in more challenging EMI conditions, such as those found near Brushless Motor Controllers. This consideration forms a part of our agenda for future development and testing, ensuring that TinyNet remains robust and reliable in a wide range of operational settings.

\subsection{Simulation}
Our software simulation of TinyNet is conducted using JavaScript on a Windows 10 laptop. The underlying code is adapted from the Simbit P2P network simulator \cite{simbit}, originally designed for blockchain technologies. We tailored this code to suit our requirements, particularly by defining the network topology as an array of arrays. Each array element represents the connections a node has on its various ports, allowing us to algorithmically model network topologies scalable to tens of thousands of nodes.

In the simulation, each processor within the network is represented as a ``manager.'' These managers are responsible for handling incoming requests on each of their ports. They maintain their own lookup and buffer depth tables and use the defined forwarding algorithm to process incoming requests. These actions are coordinated by a network controller that operates asynchronously, ensuring smooth communication across the network. Additionally, we have the capability to manually input simulation actions to these managers. This feature allows us to experiment with different communication scenarios within the network, providing a comprehensive understanding of TinyNet's behavior under various conditions.